---
title: "Programming assignment 1 (week 2)"
author: "Raphael Carvalho"
date: "18/06/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("pollutantmean.R")
source("complete.R")
source("corr.R")
```

## 1. What value is returned by the following call to pollutantmean()? You should round your output to 3 digits.

```{r 1, echo=TRUE, eval=TRUE}
pollutantmean("specdata", "sulfate", 1:10)
```

[ ] 3.782
[ ] 6.545
[ ] 4.868
[ ] 6.026
[ ] 3.666
[ x ] 4.064


## 2. What value is returned by the following call to pollutantmean()? You should round your output to 3 digits.

```{r 2, echo=TRUE, eval=TRUE}
pollutantmean("specdata", "nitrate", 70:72)
```

[ ] 2.394
[ ] 2.604
[ ] 1.182
[ ] 2.752
[ x ] 1.706
[ ] 0.914



## 3. What value is returned by the following call to pollutantmean()? You should round your output to 3 digits.

```{r 3, echo=TRUE, eval=TRUE}
pollutantmean("specdata", "sulfate", 34)
```

[ x ] 1.477
[ ] 1.300
[ ] 1.573
[ ] 0.680
[ ] 0.591
[ ] 0.450

## 4. What value is returned by the following call to pollutantmean()? You should round your output to 3 digits.

```{r 4, echo=TRUE, eval=TRUE}
pollutantmean("specdata", "nitrate")
```

[ x ] 1.703
[ ] 2.493
[ ] 1.774
[ ] 2.363
[ ] 2.233
[ ] 1.842

## 5. What value is printed at end of the following code?

```{r 5, echo=TRUE, eval=TRUE}
cc <- complete("specdata", c(6, 10, 20, 34, 100, 200, 310))
print(cc$nobs)
```

[ ] 215 201 188 204 193 213 206
[ ] 201 214 235 183 198 210 210
[ ] 204 222 200 212 213 198 196
[ ] 227 184 189 196 232 224 189
[ x ] 228 148 124 165 104 460 232
[ ] 217 210 206 214 211 203 211

## 6. What value is printed at end of the following code?

```{r 6, echo=TRUE, eval=TRUE}
cc <- complete("specdata", 54)
print(cc$nobs)
```

[ ] 213
[ x ] 219
[ ] 248
[ ] 228
[ ] 220
[ ] 205


## 7. What value is printed at end of the following code?

```{r 7, echo=TRUE, eval=TRUE}
set.seed(42)
cc <- complete("specdata", 332:1)
use <- sample(332, 10)
print(cc[use, "nobs"])

```

[ ] 608 885 684 510 765 171 244 745 624 216
[ ] 643 99 703 673 59 366 277 644 318 594
[ ] 524 577 276 487 3 592 5 148 645 435
[ ] 270 310 27 692 307 681 631 455 690 440
[ x ] 711 135 74 445 178 73 49 0 687 237

## 8. What value is printed at end of the following code?

```{r 8, echo=TRUE, eval=TRUE}
cr <- corr("specdata")                
cr <- sort(cr)                
set.seed(868)                
out <- round(cr[sample(length(cr), 5)], 4)
print(out)
```

[ ] -0.0351 0.2736 -0.0176 0.5520 0.1828
[ ] 0.3792 0.5118 0.3620 0.4726 0.5782
[ ] 0.1539 -0.0056 0.3023 0.4158 0.2558
[ ] -0.0203 0.5856 0.0983 0.3840 0.1137
[ ] 0.4474 0.4720 0.1239 0.5220 0.2538
[ x ] 0.2688 0.1127 -0.0085 0.4586 0.0447

## 9. What value is printed at end of the following code?

```{r 9, echo=TRUE, eval=TRUE}
cr <- corr("specdata", 129)                
cr <- sort(cr)                
n <- length(cr)                
set.seed(197)                
out <- c(n, round(cr[sample(n, 5)], 4))
print(out)
```

[ ] 247.0000 0.1958 0.9304 -0.4851 -0.8229 -0.0679
[ x ] 243.0000 0.2540 0.0504 -0.1462 -0.1680 0.5969
[ ] 242.0000 0.8233 0.3443 -0.2242 -0.7703 0.8735
[ ] 233.0000 -0.6377 0.3773 -0.0759 0.7335 0.2879
[ ] 225.0000 0.4216 0.4207 -0.0507 0.9377 0.0277
[ ] 229.0000 -0.2418 0.4496 0.8748 -0.3924 -0.5713

## 10. What value is printed at end of the following code?

```{r 10, echo=TRUE, eval=TRUE}
cr <- corr("specdata", 2000)                
n <- length(cr)                
cr <- corr("specdata", 1000)                
cr <- sort(cr)
print(c(n, round(cr, 4)))
```

[ ] 3.0000 0.5342 -0.6713 0.3684
[ ] 3.0000 -0.0206 -0.5881 0.5135
[ ] 3.0000 -0.8907 0.4755 -0.0175
[ ] 0.0000 -0.8974 0.8278 0.4519
[ x ] 0.0000 -0.0190 0.0419 0.1901
[ ] 2.0000 0.5596 -0.5655 -0.1241
