---
title: "Week 1 - Quiz 1"
author: "Raphael Carvalho"
date: "07/07/2019"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(xlsx)
library(rJava)
library(XML)
library(RCurl)
```


## 1.The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here: https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv and load the data into R. The code book, describing the variable names is here: https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf How many properties are worth $1,000,000 or more?


```{r q1, echo=TRUE, eval=TRUE}
fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv"
download.file(fileURL, destfile = "./quiz1_q1.csv", method="curl")

dt <- read.csv("quiz1_q1.csv")

dt %>% filter(VAL == 24) %>% summarize(n())
```

[ ] 47

[ x ] 53

[ ] 164

[ ] 31

## 2. Use the data you loaded from Question 1. Consider the variables FES in the code book. Which of the "tidy data" principles does this variable violate?


[ x ] Tidy data has one variable per column.

[ ] Each tidy data table contains information about only one type of observation.

[ ] Each variable in a tidy data set has been transformed to be interpretable.

[ ] Tidy data has one observation per row.


## 3. Download the Excel spreadsheet on Natural Gas Aquisition Program here: https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx Read rows 18-23 and columns 7-15 into R and assign the result to a variable called:

```{r q3, echo=TRUE, eval=TRUE}
fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx"
download.file(fileURL, destfile = "./quiz1_q3.xlsx", mode = "wb")
dat <- xlsx::read.xlsx("quiz1_q3.xlsx", sheetIndex = 1, rowIndex = 18:23, colIndex = 7:15)
sum(dat$Zip*dat$Ext,na.rm=T)
```

[ ] 33544718

[ x ] 36534720

[ ] 154339

[ ] NA

## 4. Read the XML data on Baltimore restaurants from here: https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml How many restaurants have zipcode 21231?

```{r q4, echo=TRUE, eval=TRUE}
fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml"
doc <- XML::xmlTreeParse(sub("s", "", fileURL), useInternal = TRUE)
rootNode <- XML::xmlRoot(doc)
zipcodes <- XML::xpathSApply(rootNode, "//zipcode", XML::xmlValue)
zipcodes <- data.table::data.table(zipcode = zipcodes)
summarize(filter(zipcodes, zipcode == "21231"), n())

```

[ x ] 127

[ ] 100

[ ] 17

[ ] 181


## 5. The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file from here: https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv using the fread() command load the data into an R object "DT". The following are ways to calculate the average of the value pwgtp15 broken down by sex. Using the data.table package, which will theliver the fastest user time?

```{r q5, echo=TRUE, eval=TRUE, warning=FALSE, error=TRUE}
DT <- data.table::fread("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv")

system.time(sapply(split(DT$pwgtp15,DT$SEX),mean))
system.time(DT[,mean(pwgtp15),by=SEX])

system.time(rowMeans(DT)[DT$SEX==1])
system.time(rowMeans(DT)[DT$SEX==2])

system.time(mean(DT$pwgtp15,by=DT$SEX))
system.time(tapply(DT$pwgtp15,DT$SEX,mean))

system.time(mean(DT[DT$SEX==1,]$pwgtp15))
system.time(mean(DT[DT$SEX==2,]$pwgtp15))
```

[ ] sapply(split(DT$pwgtp15,DT$SEX),mean)

[ x ] DT[,mean(pwgtp15),by=SEX]

[ ] rowMeans(DT)[DT$SEX==1]; rowMeans(DT)[DT$SEX==2]

[ ] mean(DT$pwgtp15,by=DT$SEX)

[ ] tapply(DT$pwgtp15,DT$SEX,mean)

[ ] mean(DT[DT$SEX==1,]$pwgtp15); mean(DT[DT$SEX==2,]$pwgtp15)