---
title: "Week 1 - Quiz 1"
author: "Raphael Carvalho"
date: "15/06/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. R was developed by statisticians working at 

[ ] The University of New South Wales

[ ] Bell Labs

[ x ] The University of Auckland

[ ] Harvard University 

## 2. The definition of free software consists of four freedoms (freedoms 0 through 3). Which of the following is NOT one of the freedoms that are part of the definition? Select all that apply. 

[ ] The freedom to redistribute copies so you can help your neighbor.

[ ] The freedom to improve the program, and release your improvements to the public, so that the whole community benefits.

[ x ] The freedom to sell the software for any price.

[ ] The freedom to run the program, for any purpose.

[ ] The freedom to study how the program works, and adapt it to your needs.

[ ] The freedom to prevent users from using the software for undesirable purposes.

[ ] The freedom to restrict access to the source code for the software.


## 3. In R the following are all atomig data types EXCEPT: (Select all that apply)

[ ] table

[ ] complex

[ x ] matrix

[ ] numeric

[ ] character

[ ] list

[ ] array

[ ] data frame

[ ] logical

[ ] integer

## 4. If I execute the expression x <- 4 in R, what is the class of the object 'x' as determined by the 'class()' function?

```{r 1, echo=TRUE, eval=TRUE}
x <- 4
class(x)
```


[ ] real

[ ] integer

[ x ] numeric

[ ] complex

[ ] vector

[ ] list

[ ] matrix

## 5. What is the class of the object defined by the expression x <- c(4, "a", TRUE)?

```{r 2, echo=TRUE, eval=TRUE}
x <- c(4, "a", TRUE)
class(x)
```

[ ] integer

[ ] numeric

[ x ] character

[ ] mixed

[ ] logical


## 6. If I have two vectors x<- c(1, 3, 5) and y <- c(3, 2, 10), what is produced by the expression cbind(x, y)?

```{r 3, echo=TRUE, eval=TRUE}
x <- c(1, 3, 5); y <- c(3, 2, 10)
cbind(x, y)
```

[ ] a vector of length 3

[ ] a 2 by 2 matrix

[ x ] a matrix with 2 columns and 3 rows

[ ] a 2 by 3 matrix

[ ] a vector of length 2

[ ] a 3 by 3 matrix

## 7. A key property of vectors in R is that 

[ ] the length of a vector must be less than 32,768

[ x ] elements of a vector all must be of the same class

[ ] a vector cannot have have attributes like dimensions

[ ] elements of a vector can only be character or numeric

[ ] elements of a vector can be of different classes

## 8. Suppose I have a list defined as x <- list(2, "a", "b", TRUE). What does x[[1]] give me? Select all that apply. 

```{r 4, echo=TRUE, eval=TRUE}
x <- list(2, "a", "b", TRUE)
x[[1]]
```

[ ] a character vector containing the element "2".

[ ] a numeric vector of length 1.

[ ] a list containing the number 2.

[ ] a list containing the letter "a".

[ x ] a numeric vector containing the element 2.

## 9. Suppose I have a vector x <- 1:4 and a vector y <- 2. What is produced by the expression x + y?

```{r 5, echo=TRUE, eval=TRUE}
x <- 1:4; y <- 2
x + y
```

[ x ] a numeric vector with elements 3, 4, 5, 6.

[ ] a numeric vector with elements 1, 2, 3, 6.

[ ] an integer vector with elements 3, 2, 3, 6.

[ ] a numeric vector with elements 3, 2, 3, 4.

[ ] a numeric vector with elements 3, 2, 3, 6.

[ ] an integer vector with elements 3, 2, 3, 4.

## 10. Suppose I have a vector x <- c(3, 5, 1, 10, 12, 6) and I want to set all elements of this vector that are less than 6 to be equal to zero. What R core achieves this? Select all that apply.

```{r 6, echo=TRUE, eval=TRUE}
# x[x >= 6] <- 0
x <- c(3, 5, 1, 10, 12, 6)
x[x >= 6] <- 0
ifelse(sum(x[c(4, 5, 6)]) == 0, TRUE, FALSE)

# x[x > 6] <- 0
x <- c(3, 5, 1, 10, 12, 6)
x[x > 6] <- 0
ifelse(sum(x[c(4, 5, 6)]) == 0, TRUE, FALSE)

# x[x != 6] <- 0
x <- c(3, 5, 1, 10, 12, 6)
x[x != 6] <- 0
ifelse(sum(x[c(4, 5, 6)]) == 0, TRUE, FALSE)

# x[x %in% 1:5] <- 0
x <- c(3, 5, 1, 10, 12, 6)
x[x %in% 1:5] <- 0
ifelse(sum(x[c(4, 5, 6)]) == 0, TRUE, FALSE)

# x[x <= 5] <- 0
x <- c(3, 5, 1, 10, 12, 6)
x[x <= 5] <- 0
ifelse(sum(x[c(4, 5, 6)]) == 0, TRUE, FALSE)

# x[x < 6] <- 0
x <- c(3, 5, 1, 10, 12, 6)
x[x < 6] <- 0
ifelse(sum(x[c(4, 5, 6)]) == 0, TRUE, FALSE)

# x[x == 6] <- 0
x <- c(3, 5, 1, 10, 12, 6)
x[x == 6] <- 0
ifelse(sum(x[c(4, 5, 6)]) == 0, TRUE, FALSE)

# x[x == 0] <- 6
x <- c(3, 5, 1, 10, 12, 6)
x[x == 0] <- 6
ifelse(sum(x[c(4, 5, 6)]) == 0, TRUE, FALSE)

# x[x < 6] == 0
x <- c(3, 5, 1, 10, 12, 6)
x[x < 6] == 0
ifelse(sum(x[c(4, 5, 6)]) == 0, TRUE, FALSE)

# x[x > 0] <- 6
x <- c(3, 5, 1, 10, 12, 6)
x[x > 0] <- 6
ifelse(sum(x[c(4, 5, 6)]) == 0, TRUE, FALSE)

# x[x == 0] < 6
x <- c(3, 5, 1, 10, 12, 6)
x[x == 0] < 6
ifelse(sum(x[c(4, 5, 6)]) == 0, TRUE, FALSE)
```

[ x ] x[x >= 6] <- 0

[ x ] x[x > 6] <- 0

[ ] x[x != 6] <- 0

[ ] x[x %in% 1:5] <- 0

[ ] x[x <= 5] <- 0

[ ] x[x < 6] <- 0

[ ] x[x == 6] <- 0

[ ] x[x == 0] <- 6

[ ] x[x < 6] == 0

[ ] x[x > 0] <- 6

[ ] x[x == 0] < 6

## 11. Use the Week 1 Quiz Data Set to answer questions 11-20.

In the dataset provided for this Quiz, what are the column names of the dataset?

```{r 7, echo=TRUE, eval=TRUE}
dt <- read.csv("hw1_data.csv")
names(dt)
```

[ x ] Ozone, Solar.R, Wind, Temp, Month, Day

[ ] Month, Day, Temp, Wind

[ ] Ozone, Solar.R, Wind

[ ] 1, 2, 3, 4, 5, 6

## 12. Extract the first 2 rows of the data frame and print them to the console. What does the output look like?

```{r 8, echo=TRUE, eval=TRUE}
head(dt, 2)
```

## 13. How manu observations (i.e. rows) are in this data frame?

```{r 9, echo=TRUE, eval=TRUE}
nrow(dt)
```

[ ] 160

[ ] 45

[ x ] 153

[ ] 129

## 14. Extract the last 2 rows of the data frame and print them to the console. What does the output look like?

```{r 10, echo=TRUE, eval=TRUE}
tail(dt, 2)
```

## 15. What is the value of Ozone in the 47th row?

```{r 11, echo=TRUE, eval=TRUE}
dt$Ozone[47]
```

[ ] 18

[ x ] 21

[ ] 63

[ ] 34

## 16. How manu missing values are in the Ozone column of this data frame?

```{r 12, echo=TRUE, eval=TRUE}
sum(is.na(dt$Ozone))
```

[ ] 9

[ ] 43

[ ] 78

[ x ] 37

## 17. What is the mean of the Ozone column in this dataset? Exclude missing values (coded as NA) from this calculation. 

```{r 13, echo=TRUE, eval=TRUE}
mean(dt$Ozone, na.rm = TRUE)
```

[ ] 18.0

[ ] 31.5

[ ] 53.2

[ x ] 42.1

## 18. Extract the subset of rows of the data frame where Ozone values are above 31 and Temp values are above 90. 

```{r 14, echo=TRUE, eval=TRUE}
dt2 <- dt[(dt$Ozone > 31 & dt$Temp>90), ]
mean(dt2$Solar.R, na.rm = TRUE)
```

[ ] 185.9

[ ] 205.0

[ ] 334.0

[ x ] 212.8

## 19. What is the mean of "Temp" when "Month" is equal to 6?

```{r 15, echo=TRUE, eval=TRUE}
dt.temp <- dt$Temp[which(dt$Month == 6)]
mean(dt.temp, na.rm = TRUE)
```

[ ] 75.3

[ ] 90.2

[ ] 85.6

[ x ] 79.1

## 20. What was the maximum ozone value in the month of May (i.e. Month is equal to 5)?

```{r 16, echo=TRUE, eval=TRUE}
dt.oz <- dt$Ozone[which(dt$Month == 5)]
max(dt.oz, na.rm = TRUE)
```

[ x ] 115

[ ] 18

[ ] 97

[ ] 100