---
title: "Week 2 - Quiz 1"
author: "Raphael Carvalho"
date: "15/06/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1. Suppose I define the following function in R: 

```{r q1_1, echo=TRUE, eval=TRUE}
cube <- function(x, n) {x^3}
```
What is the result of running
```{r q2_2, echo=TRUE, eval=TRUE}
cube(3)

```
in R after defining this function?

[ x ] The number 27 is returned

[ ] A warning is given with no value returned.

[ ] An error is returned because 'n' is not specified in the call to 'cube'

[ ] The users is prompted to specify the value of 'n'.

## 2. The following code will produce a warning in R. 
```{r q2, echo=TRUE, eval=TRUE}
x <- 1:10
if(x > 5) {
  x <- 0
}
```
Why?

[ ] There are no elements in 'x' that are greater than 5

[ x ] 'x' is a vector of length 10 and 'if' can only test a single logical statement.

[ ] You cannot set 'x' to be 0 because 'x' is a vector and 0 is a scalar.

[ ] The expression uses curly braces.

[ ] The syntax of this R expression is incorrect.


## 3. Consider the following function
```{r q3_1, echo=TRUE,eval=TRUE}
f <- function(x) {
        g <- function(y) {
                y + z
        }
        z <- 4
        x + g(x)
}
```
If I then run in R
```{r q3_2, echo=TRUE, eval=TRUE}
z <- 10
f(3)
```
What value is returned?



[ ] 4

[ ] 7

[ ] 16

[ x ] 10

## 4. Consider the following expression. What is the value of 'y' after evaluating this expression?

```{r question_4, echo=TRUE, eval=TRUE}
x <- 5
y <- if(x < 3) {
        NA
} else {
        10
}
```

[ x ] 10

[ ] NA

[ ] 5

[ ] 3

## 5. Consider the following R function. Which symbol in the above function is a free variable? 

```{r question_5_function, echo=TRUE, eval=TRUE}
h <- function(x, y = NULL, d = 3L) {
        z <- cbind(x, d)
        if(!is.null(y))
                z <- z + y
        else
                z <- z + f
        g <- x + y / z
        if(d == 3L)
                return(g)
        g <- g + 10
        g
}
```


[ x ] f

[ ] z

[ ] d

[ ] L

[ ] g

## 6. What is an enviroment in R?

[ ] a list whose elements are all functions

[ x ] a collection of symbol/value pairs

[ ] an R package that only contains data

[ ] a special type of function

## 7. The R language use what type of scoping rule for resolving free variables?

[ ] global scoping

[ ] compilation scoping

[ ] dynamic scoping

[ x ] lexical scoping

## 8. How are free variables in R functions resolved? 

[ ] The values of free variables are searched for in the environment in which the function was called

[ x ] The values of free variables are searched for in the environment in which the function was defined

[ ] The values of free variables are searched for in the working directory

[ ] The values of free variables are searched for in the global environment

## 9. What is one of the consequences of the scoping rules used in R?

[ x ] All objects must be stored in memory

[ ] All objects can be stored on the disk

[ ] Functions cannot be nested

[ ] R objects cannot be larger than 100 MB


## 10. In R, what is the parent frame?

[ ] It is the environment in which a function was defined

[ ] It is always the global environment

[ x ] It is the environment in which a function was called

[ ] It is the package search list

