---
title: 'Programming Assignment 3: Quiz'
author: "Raphael Carvalho"
date: "27/06/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
root <- "/Users/raphaelcarvalho/Documents/Cursos/JHU-datascience"
```


## 1. What result is returned by the following code?


```{r 1, echo=TRUE, eval=TRUE}
source(paste0(root, "/rprogramming/programming_assignments/best.R"))
best("SC", "heart attack")
```


[ ] LAKE CITY COMMUNITY HOSPITAL

[ ] MARY BLACK MEMORIAL HOSPITAL

[ ] CAROLINAS HOSPITAL SYSTEM

[ x ] MUSC MEDICAL CENTER

[ ] GRAND STRAND REG MED CENTER

[ ] MCLEOD MEDICAL CENTER - DARLINGTON


## 2. What result is returned by the following code?


```{r 2, echo=TRUE, eval=TRUE}
source(paste0(root, "/rprogramming/programming_assignments/best.R"))
best("NY", "pneumonia")
```


[ ] BROOKS MEMORIAL HOSPITAL

[ ] KINGSBROOK JEWISH MEDICAL CENTER

[ x ] MAIMONIDES MEDICAL CENTER

[ ] HOSPITAL FOR SPECIAL SURGERY

[ ] CROUSE HOSPITAL

[ ] NORTHERN DUTCHESS HOSPITAL

## 3. What result is returned by the following code?


```{r 3, echo=TRUE, eval=TRUE}
source(paste0(root, "/rprogramming/programming_assignments/best.R"))
best("AK", "pneumonia")
```

[ ] MT EDGECUMBE HOSPITAL

[ ] CORDOVA COMMUNITY MEDICAL CENTER

[ ] ALASKA NATIVE MEDICAL CENTER

[ x ] YUKON KUSKOKWIM DELTA REG HOSPITAL

[ ] CENTRAL PENINSULA GENERAL HOSPITAL

[ ] PROVIDENCE ALASKA MEDICAL CENTER


## 4. What result is returned by the following code?


```{r 4, echo=TRUE, eval=TRUE}
source(paste0(root, "/rprogramming/programming_assignments/rankhospital.R"))
rankhospital("NC", "heart attack", "worst")
```

[ ] CAROLINAS MEDICAL CENTER-NORTHEAST

[ x ] WAYNE MEMORIAL HOSPITAL

[ ] THOMASVILLE MEDICAL CENTER

[ ] CATAWBA VALLEY MEDICAL CENTER

[ ] SPRUCE PINE COMMUNITY HOSPITAL

[ ] WAKEMED, CARY HOSPITAL

## 5. What result is returned by the following code?


```{r 5, echo=TRUE, eval=TRUE}
source(paste0(root, "/rprogramming/programming_assignments/rankhospital.R"))
rankhospital("WA", "heart attack", 7)
```

[ ] JEFFERSON HEALTHCARE HOSPITAL

[ x ] YAKIMA VALLEY MEMORIAL HOSPITAL

[ ] TACOMA GENERAL ALLENMORE HOSPITAL

[ ] KENNEWICK GENERAL HOSPITAL

[ ] KADLEC REGIONAL MEDICAL CENTER

[ ] ST ANTHONY HOSPITAL


## 6. What result is returned by the following code?


```{r 6, echo=TRUE, eval=TRUE}
source(paste0(root, "/rprogramming/programming_assignments/rankhospital.R"))
rankhospital("TX", "pneumonia", 10)
```

[ ] UNIVERSITY OF TEXAS HEALTH SCIENCE CENTER AT TYLER

[ ] SCENIC MOUNTAIN MEDICAL CENTER

[ ] PROVIDENCE MEMORIAL HOSPITAL

[ x ] SETON SMITHVILLE REGIONAL HOSPITAL

[ ] MEMORIAL HERMANN NORTHEAST

[ ] WEST HOUSTON MEDICAL CENTER


## 7. What result is returned by the following code?


```{r 7, echo=TRUE, eval=TRUE}
source(paste0(root, "/rprogramming/programming_assignments/rankhospital.R"))
rankhospital("NY", "heart attack", 7)
```

[ ] FOREST HILLS HOSPITAL

[ ] FLUSHING HOSPITAL MEDICAL CENTER

[ ] ADIRONDACK MEDICAL CENTER

[ ] NASSAU UNIVERSITY MEDICAL CENTER

[ x ] BELLEVUE HOSPITAL CENTER

[ ] UNIVERSITY HOSPITAL S U N Y HEALTH SCIENCE CENTER



## 8. What result is returned by the following code?


```{r 8, echo=TRUE, eval=TRUE, warning=FALSE}
source(paste0(root, "/rprogramming/programming_assignments/rankall.R"))
r <- rankall("heart attack", 4)
as.character(subset(r, state == "HI")$hospital)
```

[ ] CLARK REGIONAL MEDICAL CENTER

[ x ] CASTLE MEDICAL CENTER

[ ] SANFORD USD MEDICAL CENTER

[ ] NORTH COUNTRY HOSPITAL AND HEALTH CENTER

[ ] ST FRANCIS MEDICAL CENTER

[ ] PARKLAND HEALTH AND HOSPITAL SYSTEM


## 9. What result is returned by the following code?


```{r 9, echo=TRUE, eval=TRUE, warning=FALSE}
source(paste0(root, "/rprogramming/programming_assignments/rankall.R"))
r <- rankall("pneumonia", "worst")
as.character(subset(r, state == "NJ")$hospital)
```

[ ] TERREBONNE GENERAL MEDICAL CENTER

[ ] LAURENS COUNTY HEALTHCARE SYSTEM

[ ] LAKES REGION GENERAL HOSPITAL

[ x ] BERGEN REGIONAL MEDICAL CENTER

[ ] BAY AREA HOSPITAL

[ ] POTTSTOWN MEMORIAL MEDICAL CENTER


## 10. What result is returned by the following code?


```{r 10, echo=TRUE, eval=TRUE, warning=FALSE}
source(paste0(root, "/rprogramming/programming_assignments/rankall.R"))
r <- rankall("heart failure", 10)
as.character(subset(r, state == "NV")$hospital)
```

[ ] YORK GENERAL HOSPITAL

[ x ] RENOWN SOUTH MEADOWS MEDICAL CENTER

[ ] NA

[ ] CARONDELET ST JOSEPHS HOSPITAL

[ ] CARY MEDICAL CENTER

[ ] HAWAII MEDICAL CENTER EAST
