---
title: "Subsetting Vectors"
author: "Raphael Carvalho"
date: "20/05/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Subsetting Vectors

The way you tell R that you want to select some particular elements (i.e. a 'subset') from a vector is by placing an
'index vector' in square brackets immediately following the name of the vector. For a simple example, try x[1:10] to view the first ten elements of x.

```{r 0, echo=FALSE, eval=TRUE}
x <- c(NA,NA,1.042252547,NA,NA,-0.092668232,-0.500333319,-1.404759367,NA,0.008206288,-0.330695810,-1.244128325,NA,NA,NA,-0.402158280,NA,NA,-1.494512391,-0.229679407,NA,0.278735977,-1.528015545,NA,NA,NA,-1.520262829,NA,1.610933847,0.369899514,NA,0.043526480,NA,NA,-0.054113564,-1.538931102,NA,NA,0.428470483, 0.846031063) 
```

```{r 1, echo=TRUE, eval=TRUE}
x[1:10]
```

What do you think x[is.na(x)] will give you?

```{r 2, echo=TRUE, eval=TRUE}
print("2: A vector of all NAs")
```

Recall that `!` gives us the negation of a logical expression, so !is.na(x) can be read as 'is not NA'. Therefore, if
we want to create a vector called y that contains all of the non-NA values from x, we can use y <- x[!is.na(x)]. Give
it a try.

```{r 3, echo=TRUE, eval=TRUE}
y <- x[!is.na(x)]
```

Recall that the expression y > 0 will give us a vector of logical values the same length as y, with TRUEs
corresponding to values of y that are greater than zero and FALSEs corresponding to values of y that are less than or
equal to zero. What do you think y[y > 0] will give you?

```{r 4, echo=TRUE, eval=TRUE}
print("1: A vector of all the positive elements of y")
```

 You might wonder why we didn't just start with x[x > 0] to isolate the positive elements of x. Try that now to see
why.

```{r 5, echo=TRUE, eval=TRUE}
x[x>0]
```

Since NA is not a value, but rather a placeholder for an unknown quantity, the expression NA > 0 evaluates to NA.
Hence we get a bunch of NAs mixed in with our positive numbers when we do this.


Combining our knowledge of logical operators with our new knowledge of subsetting, we could do this -- x[!is.na(x) & x > 0]. Try it out.

```{r 6, echo=TRUE, eval=TRUE}
x[!is.na(x) & x > 0]
```

Can you figure out how we'd subset the 3rd, 5th, and 7th elements of x? Hint -- Use the c() function to specify the
element numbers as a numeric vector.

```{r 7, echo=TRUE, eval=TRUE}
x[c(3, 5, 7)]
```

It's important that when using integer vectors to subset our vector x, we stick with the set of indexes {1, 2, ...,
40} since x only has 40 elements. What happens if we ask for the zeroth element of x (i.e. x[0])? Give it a try.

```{r 8, echo=TRUE, eval=TRUE}
x[0]
```

As you might expect, we get nothing useful. Unfortunately, R doesn't prevent us from doing this. What if we ask for
the 3000th element of x? Try it out.

```{r 9, echo=TRUE, eval=TRUE}
x[3000]
```

 Luckily, R accepts negative integer indexes. Whereas x[c(2, 10)] gives us ONLY the 2nd and 10th elements of x, x[c(-2,
-10)] gives us all elements of x EXCEPT for the 2nd and 10 elements.  Try x[c(-2, -10)] now to see this.

```{r 10, echo=TRUE, eval=TRUE}
x[c(-2,-10)]
```

A shorthand way of specifying multiple negative numbers is to put the negative sign out in front of the vector of
positive numbers. Type x[-c(2, 10)] to get the exact same result.

```{r 11, echo=TRUE, eval=TRUE}
x[-c(2,10)]
```

So far, we've covered three types of index vectors -- logical, positive integer, and negative integer. The only
remaining type requires us to introduce the concept of 'named' elements. Create a numeric vector with three named elements using vect <- c(foo = 11, bar = 2, norf = NA).

```{r 12, echo=TRUE, eval=TRUE}
vect <- c(foo = 11, bar = 2, norf = NA)
```

We can also get the names of vect by passing vect as an argument to the names() function. Give that a try.

```{r 13, echo=TRUE, eval=TRUE}
names(vect)
```
Alternatively, we can create an unnamed vector vect2 with c(11, 2, NA). Do that now.


```{r 14, echo=TRUE, eval=TRUE}
vect2 <- c(11, 2, NA)
```

Then, we can add the `names` attribute to vect2 after the fact with names(vect2) <- c("foo", "bar", "norf"). Go ahead.

```{r 15, echo=TRUE, eval=TRUE}
names(vect2) <- c("foo", "bar", "norf")
```

Now, let's check that vect and vect2 are the same by passing them as arguments to the identical() function.

```{r 16, echo=TRUE, eval=TRUE}
identical(vect, vect2)
```

 Now, back to the matter of subsetting a vector by named elements. Which of the following commands do you think would
give us the second element of vect?

```{r 17, echo=TRUE, eval=TRUE}
print("3: vect['bar']")
```

Likewise, we can specify a vector of names with vect[c("foo", "bar")]. Try it out.

```{r 18, echo=TRUE, eval=TRUE}
vect[c("foo", "bar")]
```
