---
title: "Workspace and Files"
author: "Raphael Carvalho"
date: "09/05/2019"
output: html_document
---

# Workspace and Files

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Determine which directory your R session is using as its current working
directory using getwd().

```{r 1, echo=TRUE, eval=TRUE}
getwd()
```


List all the objects in your local workspace using ls().

```{r 2, echo=TRUE, eval=TRUE}
ls()
```

Assign 9 to x using x <- 9.

```{r 3, echo=TRUE, eval=TRUE}
x <- 9
```

Now take a look at objects that are in your workspace using ls().

```{r 4, echo=TRUE, eval=TRUE}
ls()
```

List all the files in your working directory using list.files() or dir().

```{r 5, echo=TRUE, eval=TRUE}
x <- 9
```

List all the files in your working directory using list.files() or dir().

```{r 6, echo=TRUE, eval=TRUE}
list.files()
```

Using the args() function on a function name is also a handy way to see
what arguments a function can take. Use the args() function to determine the arguments to list.files().

```{r 7, echo=TRUE, eval=TRUE}
args(list.files)
```

Assign the value of the current working directory to a variable called
"old.dir".

```{r 8, echo=TRUE, eval=TRUE}
old.dir <- getwd()
```

Use dir.create() to create a directory in the current working directory
called "testdir".

```{r 9, echo=TRUE, eval=TRUE}
dir.create("testdir")
```

Set your working directory to "testdir" with the setwd() command.

```{r 10, echo=TRUE, eval=TRUE}
setwd("testdir")
```

Create a file in your working directory called "mytest.R" using the
file.create() function.

```{r 11, echo=TRUE, eval=TRUE}
file.create("mytest.R")
```

This should be the only file in this newly created directory. Let's check
this by listing all the files in the current directory.

```{r 12, echo=TRUE, eval=TRUE}
list.files()
```

Check to see if "mytest.R" exists in the working directory using the
file.exists() function.

```{r 13, echo=TRUE, eval=TRUE}
file.exists("mytest.R")
```

Access information about the file "mytest.R" by using file.info().

```{r 14, echo=TRUE, eval=TRUE}
file.info("mytest.R")
```

Change the name of the file "mytest.R" to "mytest2.R" by using
file.rename().

```{r 15, echo=TRUE, eval=TRUE}
file.rename("mytest.R", "mytest2.R")
```

Make a copy of "mytest2.R" called "mytest3.R" using file.copy().

```{r 16, echo=TRUE, eval=TRUE}
file.copy("mytest2.R", "mytest3.R")
```

Provide the relative path to the file "mytest3.R" by using file.path().

```{r 17, echo=TRUE, eval=TRUE}
file.path("mytest3.R")
```

You can use file.path to construct file and directory paths that are
independent of the operating system your R code is running on. Pass
'folder1' and 'folder2' as arguments to file.path to make a
platform-independent pathname.

```{r 18, echo=TRUE, eval=TRUE}
file.path("folder1", "folder2")
```

Create a directory in the current working directory called "testdir2" and a
subdirectory for it called "testdir3", all in one command by using
dir.create() and file.path().

```{r 19, echo=TRUE, eval=TRUE}
dir.create("testdir2", "testdir3", recursive = TRUE)
```

Go back to your original working directory using setwd(). (Recall that we
created the variable old.dir with the full path for the orginal working
directory at the start of these questions.)

```{r 20, echo=TRUE, eval=TRUE}
setwd(old.dir)
```

