---
title: "Subsetting Vectors"
author: "Raphael Carvalho"
date: "20/05/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Subsetting Vectors

Let's create a vector containing the numbers 1 through 20 using the `:`
operator. Store the result in a variable called my_vector.

```{r 1, echo=TRUE, eval=TRUE}
my_vector <- 1:20
```

The dim() function tells us the 'dimensions' of an object. What happens if
we do dim(my_vector)? Give it a try.

```{r 2, echo=TRUE, eval=TRUE}
dim(my_vector)
```

Clearly, that's not very helpful! Since my_vector is a vector, it doesn't
have a `dim` attribute (so it's just NULL), but we can find its length
using the length() function. Try that now.

```{r 3, echo=TRUE, eval=TRUE}
length(my_vector)
```

Ah! That's what we wanted. But, what happens if we give my_vector a `dim`
attribute? Let's give it a try. Type dim(my_vector) <- c(4, 5).

```{r 4, echo=TRUE, eval=TRUE}
dim(my_vector) <- c(4, 5)
```

It's okay if that last command seemed a little strange to you. It should!
The dim() function allows you to get OR set the `dim` attribute for an R
object. In this case, we assigned the value c(4, 5) to the `dim` attribute
of my_vector. Use dim(my_vector) to confirm that we've set the `dim` attribute correctly.

```{r 5, echo=TRUE, eval=TRUE}
dim(my_vector)
```

Another way to see this is by calling the attributes() function on
my_vector. Try it now.

```{r 6, echo=TRUE, eval=TRUE}
attributes(my_vector)
```

Just like in math class, when dealing with a 2-dimensional object (think
rectangular table), the first number is the number of rows and the second
is the number of columns. Therefore, we just gave my_vector 4 rows and 5
columns.

But, wait! That doesn't sound like a vector any more. Well, it's not. Now it's a matrix. View the contents of my_vector now to see what it looks
like.

```{r 7, echo=TRUE, eval=TRUE}
my_vector
```

Now, let's confirm it's actually a matrix by using the class() function.
Type class(my_vector) to see what I mean.

```{r 8, echo=TRUE, eval=TRUE}
class(my_vector)
```

Sure enough, my_vector is now a matrix. We should store it in a new
variable that helps us remember what it is. Store the value of my_vector in
a new variable called my_matrix.

```{r 9, echo=TRUE, eval=TRUE}
my_matrix <- my_vector
```

The example that we've used so far was meant to illustrate the point that a
matrix is simply an atomic vector with a dimension attribute. A more direct method of creating the same matrix uses the matrix() function. Now, look at the documentation for the matrix function and see if you can figure out how to create a matrix containing the same numbers (1-20) and dimensions (4 rows, 5 columns) by calling the matrix() function. Store the result in a variable called my_matrix2.

```{r 10, echo=TRUE, eval=TRUE}
my_matrix2 <- matrix(1:20, nrow = 4, ncol = 5)
```

Finally, let's confirm that my_matrix and my_matrix2 are actually identical. The identical() function will tell us if its first two arguments are the same. Try it out.

```{r 11, echo=TRUE, eval=TRUE}
identical(my_matrix, my_matrix2)
```

We may want to label the rows, so that we know which numbers belong to each
patient in the experiment. One way to do this is to add a column to the matrix, which contains the names of all four people. Let's start by creating a character vector containing the names of our patients -- Bill, Gina, Kelly, and Sean. Remember that double quotes tell R that something is a character string. Store the result in a variable called patients.


```{r 12, echo=TRUE, eval=TRUE}
patients <- c("Bill", "Gina", "Kelly", "Sean")
```

Now we'll use the cbind() function to 'combine columns'. Don't worry about storing the result in a new variable. Just call cbind() with two arguments -- the patients vector and my_matrix.

```{r 13, echo=TRUE, eval=TRUE}
cbind(patients, my_matrix)
```

Something is fishy about our result! It appears that combining the character vector with our matrix of numbers caused everything to be enclosed in double quotes. This means we're left with a matrix of character strings, which is no good.

If you remember back to the beginning of this lesson, I told you that matrices can only contain ONE class of data. Therefore, when we tried to combine a character vector with a numeric matrix, R was forced to 'coerce' the numbers to characters, hence the double quotes.

This is called 'implicit coercion', because we didn't ask for it. It just happened. But why didn't R just convert the names of our patients to numbers? I'll let you ponder that question on your own.

So, we're still left with the question of how to include the names of our patients in the table without destroying the integrity of our numeric data. Try the following -- my_data <- data.frame(patients, my_matrix)


```{r 14, echo=TRUE, eval=TRUE}
my_data <- data.frame(patients, my_matrix)
```

It looks like the data.frame() function allowed us to store our character vector of names right alongside our matrix of numbers. That's exactly what we were hoping for!

Behind the scenes, the data.frame() function takes any number of arguments and returns a single object of class `data.frame` that is composed of the original objects.Let's confirm this by calling the class() function on our newly created data frame.

```{r 15, echo=TRUE, eval=TRUE}
class(my_data)
```
Since we have six columns (including patient names), we'll need to first create a vector containing one element for each column. Create a character vector called cnames that contains the following values (in order) --z "patient", "age", "weight", "bp", "rating", "test".

```{r 16, echo=TRUE, eval=TRUE}
cnames <- c("patient", "age", "weight", "bp", "rating", "test")
```

Now, use the colnames() function to set the `colnames` attribute for our data frame. This is similar to the way we used the dim() function earlier in this lesson.

```{r 17, echo=TRUE, eval=TRUE}
colnames(my_data) <- cnames
```