---
title: "Looking at Data"
author: "Raphael Carvalho"
date: "02/06/2019"
output: pdf_document
---

Let's begin by checking the class of the plants variable with class(plants). This will give us a clue as to the overall structure of the data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)

# For compatibility with 2.2.21
.get_course_path <- function(){
  tryCatch(swirl:::swirl_courses_dir(),
           error = function(c) {file.path(find.package("swirl"),"Courses")}
  )
}

# Path to data
.datapath <- file.path(.get_course_path(),
                      'R_Programming', 'Looking_at_Data',
                      'plant-data.txt')
# Read in data
plants <- read.csv(.datapath, strip.white=TRUE, na.strings="")
```

# Looking at Data


Let's begin by checking the class of the plants variable with class(plants). This will give us a clue as to the overall structure of the data.

```{r 1, echo=TRUE, eval=TRUE}
class(plants)
```

Since the dataset is stored in a data frame, we know it is rectangular. In other words, it has two dimensions (rows and columns) and fits neatly into a table or spreadsheet. Use dim(plants) to see exactly how many rows and columns we're dealing with.

```{r 2, echo=TRUE, eval=TRUE}
dim(plants)
```


You can also use nrow(plants) to see only the number of rows. Try it out.

```{r 3, echo=TRUE, eval=TRUE}
nrow(plants)
```

... And ncol(plants) to see only the number of columns. 

```{r 4, echo=TRUE, eval=TRUE}
ncol(plants)
```

If you are curious as to how much space the dataset is occupying in memory, you can use object.size(plants).

```{r 5, echo=TRUE, eval=TRUE}
object.size(plants)
```

Now that we have a sense of the shape and size of the dataset, let's get a feel for what's inside. names(plants) will return a character vector of column (i.e. variable) names. Give it a shot.

```{r 6, echo=TRUE, eval=TRUE}
names(plants)
```

We've applied fairly descriptive variable names to this dataset, but that won't always be the case. A logical next step is to peek at the actual data. However, our dataset contains over 5000 observations (rows), so it's impractical to view the whole thing all at once.

The head() function allows you to preview the top of the dataset. Give it a try with only one argument.

```{r 7, echo=TRUE, eval=TRUE}
head(plants)
```

By default, head() shows you the first six rows of the data. You can alter this behavior by passing as a second argument the number of rows you'd like to view. Use head() to preview the first 10 rows of plants.


```{r 8, echo=TRUE, eval=TRUE}
head(plants, 10)
```

The same applies for using tail() to preview the end of the dataset. Use tail() to view the last 15 rows.

```{r 9, echo=TRUE, eval=TRUE}
tail(plants, 15)
```

After previewing the top and bottom of the data, you probably noticed lots of NAs, which are R's placeholders for missing values. Use summary(plants) to get a better feel for how each variable is distributed and how much of the dataset is missing.

```{r 10, echo=TRUE, eval=TRUE}
summary(plants)
```

You can see that R truncated the summary for Active_Growth_Period by including a catch-all category called 'Other'. Since it is a categorical/factor variable, we can see how many times each value actually occurs in the data with table(plants$Active_Growth_Period).

```{r 11, echo=TRUE, eval=TRUE}
table(plants$Active_Growth_Period)
```

Perhaps the most useful and concise function for understanding the *str*ucture of your data is str(). Give it a try now.

```{r 12, echo=TRUE, eval=TRUE}
str(plants)
```