---
title: 'swirl Lesson 1: Basic Building Blocks'
author: "Raphael Carvalho"
date: "08/05/2019"
output: html_document
---
# Basic Building blocks 

* In its simplest form, R can be used as an interactive calculator. Type 5 +
7 and press Enter.

```{r 1, echo=TRUE, eval=TRUE}
5 + 7
```

* To assign the result of 5 + 7 to a new variable called x, you type x <- 5 + 7. This can be read as 'x gets 5 plus 7'. Give it a try now.

```{r 2, echo=TRUE, eval=TRUE}
x <- 5 + 7
```

* To view the contents of the variable x, just type x and press Enter. Try it
now.

```{r 3, echo=TRUE, eval=TRUE}
x
```

* Now, store the result of x - 3 in a new variable called y.

```{r 4, echo=TRUE, eval=TRUE}
y <- x - 3
```

* What is the value of y? Type y to find out.

```{r 5, echo=TRUE, eval=TRUE}
y
```

* The easiest way to create a vector is with the c() function, which stands
for 'concatenate' or 'combine'. To create a vector containing the numbers
1.1, 9, and 3.14, type c(1.1, 9, 3.14). Try it now and store the result in
a variable called z

```{r 6, echo=TRUE, eval=TRUE}
z <- c(1.1, 9, 3.14)
```

* Anytime you have questions about a particular function, you can access R's
built-in help files via the `?` command. For example, if you want more
information on the c() function, type ?c without the parentheses that
normally follow a function name. Give it a try.

```{r 7, echo=TRUE, eval=TRUE}
?c
```

* Type z to view its contents. Notice that there are no commas separating the values in the output.

```{r 8, echo=TRUE, eval=TRUE}
z
```

* You can combine vectors to make a new vector. Create a new vector that
contains z, 555, then z again in that order. Don't assign this vector to a
new variable, so that we can just see the result immediately.

```{r 9, echo=TRUE, eval=TRUE}
c(z, 555, z)
```

* Numeric vectors can be used in arithmetic expressions. Type the following
to see what happens: z * 2 + 100.

```{r 10, echo=TRUE, eval=TRUE}
z * 2 + 100
```

* Take the square root of z - 1 and assign it to a new variable called
my_sqrt.

```{r 11, echo=TRUE, eval=TRUE}
my_sqrt <- sqrt(z-1)
```

* Before we view the contents of the my_sqrt variable, what do you think it
contains?

*2: a vector of length 3*

* Print the contents of my_sqrt.

```{r 12, echo=TRUE, eval=TRUE}
my_sqrt
```

* Now, create a new variable called my_div that gets the value of z divided
by my_sqrt.

```{r 13, echo=TRUE, eval=TRUE}
my_div <- z/my_sqrt
```

* Which statement do you think is true?

*3: The first element of my_div is equal to the first element of z divided by the first element of my_sqrt, and so on...*

* Go ahead and print the contents of my_div.

```{r 14, echo=TRUE, eval=TRUE}
my_div
```

* To see another example of how this vector 'recycling' works, try adding
c(1, 2, 3, 4) and c(0, 10). Don't worry about saving the result in a new
variable.

```{r 15, echo=TRUE, eval=TRUE}
c(1, 2, 3, 4) + c(0, 10)
```

* Try c(1, 2, 3, 4) + c(0, 10, 100) for an example.

```{r 16, echo=TRUE, eval=TRUE}
 c(1, 2, 3, 4) + c(0, 10, 100)
```

* In many programming environments, the up arrow will cycle through previous
commands. Try hitting the up arrow on your keyboard until you get to this
command (z * 2 + 100), then change 100 to 1000 and hit Enter. If the up
arrow doesn't work for you, just type the corrected command.

```{r 17, echo=TRUE, eval=TRUE}
z * 2 + 1000
```

* You can type the first two letters of the variable name, then hit the Tab
key (possibly more than once). Most programming environments will provide a
list of variables that you've created that begin with 'my'. This is called
auto-completion and can be quite handy when you have many variables in your
workspace. Give it a try. (If auto-completion doesn't work for you, just
type my_div and press Enter.)

```{r 18, echo=TRUE, eval=TRUE}
my_div
```