---
title: "Sequence of Numbers"
author: "Raphael Carvalho"
date: "09/05/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Missing Values

Any operation involving NA generally yields NA as the result. To illustrate, let's create a vector c(44, NA, 5, NA)
and assign it to a variable x.

```{r 1, echo=TRUE, eval=TRUE}
x <- c(44, NA, 5, NA)
```

Now, let's multiply x by 3.

```{r 2, echo=TRUE, eval=TRUE}
x * 3
```

To make things a little more interesting, lets create a vector containing 1000 draws from a standard normal
distribution with y <- rnorm(1000).

```{r 3, echo=TRUE, eval=TRUE}
y <- rnorm(1000)
```

Next, let's create a vector containing 1000 NAs with z <- rep(NA, 1000).

```{r 4, echo=TRUE, eval=TRUE}
z <- rep(NA, 1000)
```

Finally, let's select 100 elements at random from these 2000 values (combining y and z) such that we don't know how
many NAs we'll wind up with or what positions they'll occupy in our final vector -- my_data <- sample(c(y, z), 100).

```{r 5, echo=TRUE, eval=TRUE}
my_data <- sample(c(y, z), 100)
```

Let's first ask the question of where our NAs are located in our data. The is.na() function tells us whether each
element of a vector is NA. Call is.na() on my_data and assign the result to my_na.

```{r 6, echo=TRUE, eval=TRUE}
my_na <- is.na(my_data)
```

In our previous discussion of logical operators, we introduced the `==` operator as a method of testing for equality
between two objects. So, you might think the expression my_data == NA yields the same results as is.na(). Give it a
try.

```{r 7, echo=TRUE, eval=TRUE}
my_data == NA
```

Let's give that a try here. Call the sum() function on my_na to count the total number of TRUEs in my_na, and thus the
total number of NAs in my_data. Don't assign the result to a new variable.

```{r 8, echo=TRUE, eval=TRUE}
sum(my_na)
```

Now that we've got NAs down pat, let's look at a second type of missing value -- NaN, which stands for 'not a number'.
To generate NaN, try dividing (using a forward slash) 0 by 0 now.

```{r 9, echo=TRUE, eval=TRUE}
0/0
```

Let's do one more, just for fun. In R, Inf stands for infinity. What happens if you subtract Inf from Inf?

```{r 10, echo=TRUE, eval=TRUE}
Inf/Inf
```

