---
title: "Vectors"
author: "Raphael Carvalho"
date: "15/05/2019"
output: html_document
---

# Vectors

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, create a numeric vector num_vect that contains the values 0.5, 55, -10, and 6.

```{r 1, echo=TRUE, eval=TRUE}
num_vect <- c(0.5, 55, -10, 6)
```

Now, create a variable called tf that gets the result of num_vect < 1, which is read
as 'num_vect is less than 1'.

```{r 2, echo=TRUE, eval=TRUE}
tf <- num_vect < 1
```

What do you think tf will look like?

```{r 3, echo=TRUE, eval=TRUE}
print("a vector of 4 logical values")
```

Let's try another. Type num_vect >= 6 without assigning the result to a new variable.

```{r 4, echo=TRUE, eval=TRUE}
num_vect >= 6
```


(3 > 5) & (4 == 4)

```{r 5, echo=TRUE, eval=TRUE}
print(1: FALSE)
```

(TRUE == TRUE) | (TRUE == FALSE)

```{r 6, echo=TRUE, eval=TRUE}
print(2: TRUE)
```

((111 >= 111) | !(TRUE)) & ((4 + 1) == 5)

```{r 7, echo=TRUE, eval=TRUE}
print(2: TRUE)
```

Create a character vector that contains the following words: "My", "name", "is". Remember to enclose each word in its own set of double quotes, so that R knows they
are character strings. Store the vector in a variable called my_char.

```{r 8, echo=TRUE, eval=TRUE}
my_char <- c("My", "name", "is")
```

Type paste(my_char, collapse = " ") now. Make sure there's a space between the double
quotes in the `collapse` argument. You'll see why in a second.

```{r 9, echo=TRUE, eval=TRUE}
paste(my_char, collapse = " ") 
```

To add (or 'concatenate') your name to the end of my_char, use the c() function like
this: c(my_char, "your_name_here"). Place your name in double quotes where I've put
"your_name_here". Try it now, storing the result in a new variable called my_name.

```{r 10, echo=TRUE, eval=TRUE}
my_name <- c(my_char, "Raphael")
```
